\chapter{Related Work}
\label{chap:related}

The compiler presented in this report is unusual in that is takes into its front-end a bytecode file, one that has previously been compiled by another program. There is currently no pre-existing Dalvik executable front-end for LLVM, and LLVM's Java front-end is stale and unsupported\footnotemark \footnotetext[1]{http://old.nabble.com/Re\%3A-compiling-java-frontend-p25500890.html}. The code for this project had to be started from scratch.

LLVM offers a sub-project, VMKit, a framework for building virtual machines \cite{geoffray10vmkit}. VMKit provides a Java virtual machine on top of LLVM to run Java applications. There is also partial support for ahead-of-time compilation from Java code to native code. Another of LLVM's sub-projects, dragonegg \cite{dragonegg}, integrates LLVM optimisers and code generator with GCC 4.5 parsers. It is currently able to compile simple Java programs, but they are unable to execute properly: arising because GCC's link-time optimisation is not supported by the Java front-end. This is, however, the closest that the LLVM project has to a usable Java front-end.

The majority of existing compilers that work with bytecode as input are just-in-time compilers, whereas the compiler detailed in this report is a \emph{static} bytecode compiler. One such example of a static bytecode compiler is Caffeine\cite{Hsieh:1996:JBN:243846.243864}, an optimising translator from Java bytecode to native machine code. Hsieh et al. list the difficulties involved in translating from a stack-based programming platform like Java bytecode to a register-based machine. The Dalvik virtual machine, unlike its Java-based relative, is a register-based platform, and so in the case of the Dalvik-to-LLVM compiler, this challenge is not present.

In the same vein as Caffeine is Soot \cite{Vallee-Rai:1999:SJB:781995.782008}, a framework for optimising Java bytecode. This project translates Java bytecode to further-optimised Java bytecode. This is a similar approach to our compiler in that both its input and output are intermediate representations. Soot optimises Java bytecode by introducing three intermediate representations: Baf, Jimple, and Grimp. The process of optimisation is to transform the Java bytecode to Baf, Jimple, Grimp, back to Baf, then to bytecode, while at each stage performing an appropriate optimisation.

The results presented by Soot show that there is a call for the optimisation of Java bytecode. Soot shows up to 8\% improvement when the optimized bytecode is run using the interpreter and up to 21\% when run using the JIT compiler. This indicates that the optimisations offered by LLVM can too help improve runtime performance of Dalvik executables. In order to be processable by the LLVM analyser and optimiser, they first need to be converted to LLVM bytecode. This provides motivation for the compiler presented in this report.


In addition to the previous two compilers is the Jalape\~{n}o Dynamic Optimising Compiler \cite{Burke:1999:JDO:304065.304113}. This paper presents a compiler for the Jalape\~{n}o virtual machine, a new Java virtual machine by the same group. As opposed to other Java virtual machines which provide a just-in-time compiler and an interpreter, the Jalape\~{n}o virtual machine always translates bytecodes to native code prior to execution. Similar to the previous projects, Jalape\~{n}o takes Java bytecode as input and translates it statically to native machine code.

The Jalape\~{n}o Dynamic Optimising Compiler was evaluated against the IBM enhanced port of the Sun JDK 1.1.16, and reported speedups in three of nine selected benchmarks.


Of particular note is Swift\cite{Zhang:2012:SRJ:2151024.2151035}. Swift is a register-based JIT compiler for embedded JVMs. The Swift prototype presented in the paper is implemented upon the Dalvik executable format. Swift achieves impressive performance results. Swift achieves a speedup of 3.13 over the best-performing interpreter on the selected benchmarks. Compared with the state-of-the-art just-in-time compiler in Android, JITC-Droid, Swift achieves a speedup of 1.42.

Furthermore, compilers that work in translating from one programming \emph{paradigm} to another have been around for years. In the case of the translated from the Dalvik executable format to LLVM bytecode, this involves a conversion from an object-oriented paradigm to a procedural one. This is a well-understood problem, as almost all object-oriented languages must be compiled down to non-object-oriented code at some point.

Indeed, LLVM can be used to convert from object-oriented C++ to procedural C\footnotemark \footnotetext[2]{http://llvm.org/docs/FAQ.html\#translatecxx}. There are some limitations to this approach, namely, it does not eliminate the need for C++ library support. This is a similar case to the one encountered during the translation from Dalvik to LLVM: a standard Java library implementation is still necessary to execute the compiled programs. LLVM's C back-end also does not support C++-ABI-conforming exceptions.

Cfont was the original C++ compiler, developed by the creator of C++: Bjarne Stroustrup \cite{Stroustrup:2007:ELR:1238844.1238848}. Instead of compiling C++ to native code, it instead converted C++ code to C. The project was abandoned in 1993 after a failed attempt to add exception support.

Other compilers that can compile C++ to C include EDG's C++ to C translator\footnotemark \footnotetext[3]{http://www.edg.com}, and the one offered by Comeau Computing\footnotemark \footnotetext[4]{http://www.comeaucomputing.com}.

While there are no existing direct implementations of the compiler presented in this report, there are several projects working in a similar fashion. The compiler presented in this report is unusual in that it statically translates from one machine-generated language to another. As seen by Caffeine, Soot, and the Jalape\~{n}o compiler, however, there do exist other compilers which statically-compile bytecode to another (or the same) language. In addition, this Dalvik-to-LLVM compiler must translate an object-oriented language down to a low-level, procedural one. The task can be achieved using LLVM, Cfront, or a number of alternatives, as demonstrated.
